---
title: "Assignment5"
author: "Bryce Owen, Abbie McKay, and Kimberly Tang"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Import, Read, and Preview Data

```{r}
library("readxl")
library("tidyverse")
library(ggplot2)
buildings <- read_excel('tallestbuildings.xlsx')
buildings
```

### Rename columns and clean up data

```{r}
library(rattle) # need to install the package first
library(magrittr) # need to install that package first
names(buildings) %<>% normVarNames()
buildings
```

### Group the buildings by City in descending order

```{r}
Citycounts=buildings%>%
  group_by(city)%>%
  summarise(number=length(city))
arrange(Citycounts, desc(number))
```

### Group and rank means of city buildings and plot

```{r}
Citycounts=buildings%>%
  group_by(city)%>%
  summarise(mean_height=mean(height_ft))
mean_buildings_sort <- arrange(Citycounts, desc(mean_height))

ggplot(data = mean_buildings_sort, mapping = aes(reorder(city, mean_height), mean_height)) + geom_bar(stat = 'identity') + coord_flip() + ggtitle('Mean Height of Tallest Buildings by City') + labs(y = 'Mean of Building Heights (ft)', x = 'City')
```

### Create new columns by separating the city and country

```{r}
library(countrycode)
buildings <- buildings %>%
  separate(city, into = c('city', 'country'), sep = '[:punct:]')
```

### Group the buildings by Country in descending order

```{r}
# use the country code function
buildings$country = countrycode(buildings$country, "iso2c", "country.name")
Countrycounts=buildings%>%
  group_by(country)%>%
  summarise(number=length(country))
arrange(Countrycounts, desc(number))
```

### Group and rank means of country buildings and plot

```{r}
Countrycounts=buildings%>%
  group_by(country)%>%
  summarise(mean_height=mean(height_ft))
mean_buildings_sort <- arrange(Countrycounts, desc(mean_height))

ggplot(data = mean_buildings_sort, mapping = aes(reorder(country, mean_height), mean_height)) + geom_bar(stat = 'identity') + coord_flip() + ggtitle('Mean Height of Tallest Buildings by Country') + labs(y = 'Mean of Building Heights (ft)', x = 'Country')
```

### Color each bar for a country based on the number of buildings from this dataset that are present in that country

```{r}
Countrycounts=buildings%>%
  group_by(country)%>%
  summarise(mean_height=mean(height_ft), count=length(country))
Countrycounts
ggplot(data = Countrycounts, mapping = aes(reorder(country, mean_height), mean_height, fill = as.factor(count))) + geom_bar(stat = 'identity') + coord_flip() + ggtitle('Mean Height of Tallest Buildings by Country') + labs(y = 'Mean of Building Heights (ft)', x = 'Country') + guides(fill=guide_legend(title="Building Count"))
```

### Split use of buildings, pivot longer, and comparing mean heights for building purposes

```{r}
buildings <- buildings %>%
  separate(use, into = c('use1', 'use2', 'use3'), sep = ' / ')
BuildingPurpose=buildings%>%
  # Extra non-essential step to visualize df progression and split
  group_by(building_name)%>%
  # Extra non-essential step to visualize df progression and split
  summarise(use1, use2, use3, height_ft)%>%
  pivot_longer(cols = c('use1', 'use2', 'use3'), values_to = "use", values_drop_na = TRUE)

BuildingPurpose%>%
  group_by(use)%>%
  summarise(mean_height=mean(height_ft))
```

